---
title: "HCC early recurrence model"
author: 
  - En Hu
  
documentclass: ctexart
geometry: margin=0.6cm
output:
  rticles::ctex:
    fig_caption: yes
    number_sections: no
    toc: yes
header-includes:
 - \usepackage{fvextra}
 - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
 - \CTEXsetup[format={\huge\bfseries\raggedright}]{section}
---
# Data description
1. all_data: MVI status (0:present,1:absent), GSN concentrations (ug/ul), AFP status (0:negative,1:positive) and DCP status (0:negative,1:positive) group (0:recurrence time ≥12m, 1:recurrence time <12m)
2. df_cut:  the results of 10 times 5-flod cross modelling are stored, the first column is seed+fold (5661: seed=566,fold=1),the second column stores the response of the model and the third column is the real label.

# 1. 5-fold validation
```{r}
aucs <- c()
sens <- c()
spec <- c()
roclist <- list()
df_cut <- data.frame()
cutoff <- c()

set.seed(2008)
for(i in sample(1:1000,size=10)){
  for(t in 1:5){
    set.seed(i)
    folds <- createFolds(y=all_data$group,k=5)
    train_cv <- all_data[-folds[[t]],]
    test_cv <- all_data[folds[[t]],]
    pre <- glm(group~GSN+MVI,family=binomial(link = "logit"),data = train_cv)
    auc <- pROC::roc(test_cv$group,predict(pre,test_cv,type="response"))$auc %>% as.numeric() %>% round(.,2)
    coor <- coords(pROC::roc(test_cv$group,predict(pre,test_cv,type="response")),"best")
    cutoff <- c(cutoff,coor[,1])
    sens<-c(sens,coor[,3])
    spec<-c(spec,coor[,2])
    roclist[[paste0(i,t)]] <- pROC::roc(test_cv$group,predict(pre,test_cv,type="response"))
    aucs <- c(aucs,auc)
    df_cut <- rbind(df_cut,data.frame(name=paste0(i,t),pre=predict(pre,test_cv,type="response"),group=test_cv$group))
  }
}
```
# 2. Figure 6B
## Function description
1. mean_roc: Calculate average roc based on 50 modelling results
```{r}
mean_roc <- function(data, cutoffs = seq(from = 0, to = 1, by = 0.05)) {
    map_df(cutoffs, function(cp) {
        out <- cutpointr(data = data, 
                         x = pre, 
                         class = group, 
                         subgroup = name, 
                         method = oc_manual, 
                         cutpoint = cp,
                         pos_class = 1,
                         direction = ">=")
        data.frame(cutoff = cp,
                   sensitivity = mean(out$sensitivity),
                   specificity = mean(out$specificity))
    })
}


roc_data <- ggplot_build(ggroc(roclist))
roc_data <- roc_data$plot$data
mr <- mean_roc(df_cut)
CI(aucs)# calculate aucs
p2 <- data.frame(threshold=0,specificity=mr$specificity,sensitivity=mr$sensitivity,name="mean") %>%
    rbind(roc_data) %>%
    mutate(name2=ifelse(name=="mean","mean","other")) %>%
    
    ggplot(.)+
    # geom_ribbon(data=mr3,aes(ymin=lower,ymax=upper,x=1-x))+
    # geom_ribbon(data=mr4,aes(xmin=lower,xmax=upper,y=y))+
    geom_step(aes(group=name,color=name2,linetype=name2,x=1-specificity,y=sensitivity),size=1) +
    scale_color_manual(values = c("red","grey90"))+
    theme_classic()+
    
    geom_segment(x=0,y=0,xend=1,yend=1,linetype=2,size=1,color="gray20")+

    # scale_x_continuous(expand = c(0,0))+
    # scale_y_continuous(expand = c(0,0))+
    geom_segment(x=-Inf,xend=Inf,y=Inf,yend=Inf,linewidth=rel(1.2))+
    geom_segment(x=Inf,xend=Inf,y=-Inf,yend=Inf,linewidth=rel(1.2))+
    
    theme(axis.line = element_line(linewidth = rel(1.5)),
          axis.ticks = element_line(colour = "black",size=1),
          axis.text = element_text(family = "TNR",face = "bold",color = "black",size = 15),
          axis.title = element_text(family = "TNR",face = "bold",color = "black",size = 18),
          legend.text = element_text(family = "TNR",face = "bold",color = "black",size = 12),
          #legend.title = element_text(family = "TNR",face = "bold",color = "black",size = 15),
          legend.title = element_blank(),
          legend.position = "none")+
    geom_text(x=0.5,y=0.4,label="AUC(95% CI):0.803 [0.820,0.783]",size=5)+
    labs(x="1-Specificity",y="Sensitivity")


cairo_pdf(filename = "./Out/5fold_roc.pdf",height = 5.02,width = 4.96)
p2
dev.off()
```
# 3. One ROC closer to the average ROC curve was selected for further evaluation of the model
```{r}
set.seed(749)
folds <- createFolds(y=all_data$group,k=5)
train_cv <- all_data[-folds[[4]],]
test_cv <- all_data[folds[[4]],]
pre <- glm(group~GSN+MVI,family=binomial(link = "logit"),data = train_cv)
summary(pre)
pROC::roc(test_cv$group,predict(pre,test_cv,type="response"))
```

# 4. nomogram:Figure 6F
```{r}
logistic_lrm <- lrm(group~GSN+MVI,data = train_cv,x = T,y = T)
dd <- datadist(train_cv)
options(datadist="dd")
nom <- nomogram(fit = logistic_lrm,fun = function(x)1/(1+exp(-x)),lp=T,funlabel = "12m Recurrence risk ")
cairo_pdf(filename = "./Out/nom.pdf",height = 8.10,width = 8.10)
plot(nom)
dev.off()
```
# 5. figure 6C
```{r}
roc11 <- pROC::roc(train_cv$group,predict(pre,train_cv,type="response"))
roc22 <- pROC::roc(train_cv$group,train_cv$MVI)
roc33 <- pROC::roc(train_cv$group,train_cv$GSN)
roc44 <- pROC::roc(train_cv$group,train_cv$DCP)
roc55 <- pROC::roc(train_cv$group,train_cv$AFP)
  x2 <-  ggroc(list(Model=roc11,
                  MVI=roc22,
                  GSN=roc33,
                  DCP=roc44,
                  AFP=roc55))
 df2 <- x2$data
 df2$specificity <- 1-df2$specificity
 df2$name <- factor(df2$name,levels = c("Model","GSN","MVI","AFP","DCP"),
                   labels = c("Mdoel:0.805","GSN:0.705","MVI:0.693","AFP:0.614","DCP:0.569"))
 
p5 <- ggplot(df2,aes(x=specificity,y=sensitivity))+
     geom_segment(x=0,y=0,xend=1,yend=1,linetype=2,size=1,color="gray40")+
     geom_line(aes(group=name,color=name),size=1)+
     theme_classic()+
     # scale_x_continuous(expand = c(0,0))+
     # scale_y_continuous(expand = c(0,0))+
     geom_segment(x=-Inf,xend=Inf,y=Inf,yend=Inf,linewidth=rel(1.2))+
     geom_segment(x=Inf,xend=Inf,y=-Inf,yend=Inf,linewidth=rel(1.2))+
     theme(axis.line = element_line(linewidth = rel(1.5)),
           axis.ticks = element_line(colour = "black",size=1),
           axis.text = element_text(family = "TNR",face = "bold",color = "black",size = 15),
           axis.title = element_text(family = "TNR",face = "bold",color = "black",size = 18),
           legend.text = element_text(family = "TNR",face = "bold",color = "black",size = 12),
           #legend.title = element_text(family = "TNR",face = "bold",color = "black",size = 15),
           legend.title = element_blank(),
           legend.position = c(0.8,0.25))+
     scale_color_manual(values = c("#D62728FF","#1F77B4FF" ,"#FF7F0EFF" ,"#2CA02CFF" ,"#9467BDFF"))+
     labs(x="1-Specificity",y="Sensitivity")
 
 
cairo_pdf(filename = "./Out/ROC_trian.pdf",height = 4.8,width = 4.9)
p5
dev.off()
```
# 6. figur 6D
```{r}
roc111 <- pROC::roc(test_cv$group,predict(pre,test_cv,type="response"));roc111
roc222 <- pROC::roc(test_cv$group,test_cv$MVI);roc222
roc333 <- pROC::roc(test_cv$group,test_cv$GSN);roc333
roc444<- pROC::roc(test_cv$group,test_cv$DCP);roc444
roc555 <- pROC::roc(test_cv$group,test_cv$AFP);roc555

  x3 <-  ggroc(list(Model=roc111,
                  MVI=roc222,
                  GSN=roc333,
                  DCP=roc444,
                  AFP=roc555))

df3 <- x3$data
 df3$specificity <- 1-df3$specificity
 df3$name <- factor(df3$name,levels = c("Model","GSN","MVI","AFP","DCP"),
                   labels = c("Mdoel:0.805","GSN:0.754","MVI:0.587","AFP:0.574","DCP:0.621"))
 
p6 <- ggplot(df3,aes(x=specificity,y=sensitivity))+
     geom_segment(x=0,y=0,xend=1,yend=1,linetype=2,size=1,color="gray40")+
     geom_line(aes(group=name,color=name),size=1)+
     theme_classic()+
     # scale_x_continuous(expand = c(0,0))+
     # scale_y_continuous(expand = c(0,0))+
     geom_segment(x=-Inf,xend=Inf,y=Inf,yend=Inf,linewidth=rel(1.2))+
     geom_segment(x=Inf,xend=Inf,y=-Inf,yend=Inf,linewidth=rel(1.2))+
     theme(axis.line = element_line(linewidth = rel(1.5)),
           axis.ticks = element_line(colour = "black",size=1),
           axis.text = element_text(family = "TNR",face = "bold",color = "black",size = 15),
           axis.title = element_text(family = "TNR",face = "bold",color = "black",size = 18),
           legend.text = element_text(family = "TNR",face = "bold",color = "black",size = 12),
           #legend.title = element_text(family = "TNR",face = "bold",color = "black",size = 15),
           legend.title = element_blank(),
           legend.position = c(0.8,0.25))+
     scale_color_manual(values = c("#D62728FF","#1F77B4FF" ,"#FF7F0EFF" ,"#2CA02CFF" ,"#9467BDFF"))+
     labs(x="1-Specificity",y="Sensitivity")
 
 
cairo_pdf(filename = "./Out/ROC_valid.pdf",height = 4.8,width = 4.9)
p6
dev.off()
```
# 7. figure 6E
```{r}
all_data2 <- all_data[all_data$AFPPositive==0,]
roc1111 <- pROC::roc(dus2_n$group,predict(pre,dus2_n,type="response"));roc1111
roc2222 <- pROC::roc(dus2_n$group,dus2_n$MVI);roc2222
roc3333 <- pROC::roc(dus2_n$group,dus2_n$GSN);roc3333
roc4444<- pROC::roc(dus2_n$group,dus2_n$DCP);roc4444

x4 <-  ggroc(list(Model=roc1111,
                  MVI=roc2222,
                  GSN=roc3333,
                  DCP=roc4444))

df4 <- x4$data
 df4$specificity <- 1-df4$specificity
 df4$name <- factor(df4$name,levels = c("Model","GSN","MVI","DCP"),
                   labels = c("Mdoel:0.798","GSN:0.675","MVI:0.672","DCP:0.600"))
 
p7 <- ggplot(df4,aes(x=specificity,y=sensitivity))+
     geom_segment(x=0,y=0,xend=1,yend=1,linetype=2,size=1,color="gray40")+
     geom_line(aes(group=name,color=name),size=1)+
     theme_classic()+
     # scale_x_continuous(expand = c(0,0))+
     # scale_y_continuous(expand = c(0,0))+
     geom_segment(x=-Inf,xend=Inf,y=Inf,yend=Inf,linewidth=rel(1.2))+
     geom_segment(x=Inf,xend=Inf,y=-Inf,yend=Inf,linewidth=rel(1.2))+
     theme(axis.line = element_line(linewidth = rel(1.5)),
           axis.ticks = element_line(colour = "black",size=1),
           axis.text = element_text(family = "TNR",face = "bold",color = "black",size = 15),
           axis.title = element_text(family = "TNR",face = "bold",color = "black",size = 18),
           legend.text = element_text(family = "TNR",face = "bold",color = "black",size = 12),
           #legend.title = element_text(family = "TNR",face = "bold",color = "black",size = 15),
           legend.title = element_blank(),
           legend.position = c(0.8,0.25))+
     scale_color_manual(values = c("#D62728FF","#1F77B4FF" ,"#FF7F0EFF" ,"#2CA02CFF" ,"#9467BDFF"))+
     labs(x="1-Specificity",y="Sensitivity")
 
 
cairo_pdf(filename = "./Out/ROC_afpn.pdf",height = 4.8,width = 4.9)
p7
dev.off() 

```
